---
title: "SoundCloud Analysis"
author: Hajara Muzammal
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    toc: true
    toc-depth: 3
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

# **Introduction:**

Over the past several decades, the consumption and discovery of music
have undergone a dramatic transformation, shifting from physical media
and radio programming toward digital streaming platforms powered by
algorithms. Modern listeners rarely depend on CDs, radio countdowns, or
store recommendations to find new music. Instead, they increasingly rely
on automated systems such as curated playlists, algorithmic feeds, and
social recommendations. Major platforms like Spotify leverage machine
learning models to anticipate listener taste and predict song
preferences, while community-based platforms like SoundCloud rely
heavily on reposts and follower-driven exposure. Together, these
features decentralize distribution, making pathways to musical success
less linear and less controlled by industry gatekeepers than in previous
eras.

In this emerging landscape, playlists play a particularly influential
role. Playlists serve as curated gateways through which music is
discovered, circulated, and eventually popularized. Many believe that
being featured in a widely followed playlist directly contributes to a
song’s rise in popularity. However, there is a gap between public
perception and data-based evidence on whether playlist inclusion
realistically drives visibility and streaming success — or whether
popularity is largely shaped by external social dynamics such as viral
sharing, artist branding, and digital marketing investment.

Motivated by this uncertainty, this report analyzes nearly 15,000 tracks
from a SoundCloud-linked dataset hosted on Hugging Face, which combines
playlist metadata, Spotify-style audio features, track popularity
scores, and direct streaming links. This provides a rare opportunity to
explore streaming metrics from both algorithmic and creator-driven
perspectives. The guiding research question is therefore: **Is track
popularity associated with playlist inclusion?** A supporting objective
considers whether measurable audio characteristics contribute to
popularity outcomes. By applying exploratory visuals and statistical
methods, this analysis aims to determine whether songs that appear in
more playlists or songs with certain musical traits tend to rise to
greater levels of popularity — or whether these assumptions lack
empirical support.

# **Data Ingest**

We use a publicly available dataset hosted on Hugging Face, which
contains playlist metadata, song characteristics, and direct SoundCloud
links.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

url <- "https://huggingface.co/datasets/Zuru7/Spotify_Songs_with_SoundCloud_links/resolve/main/song_df_normalised.csv"
SONGS_raw <- read_csv(url, show_col_types = FALSE)

# Standardize names (works even if you re-run the doc)
SONGS <- SONGS_raw %>%
  rename(
    track          = any_of(c("track", "track_name")),
    artist         = any_of(c("artist", "track_artist")),
    album          = any_of(c("album", "track_album_name")),
    popularity     = any_of(c("popularity", "track_popularity")),
    playlist_genre = any_of(c("genre", "playlist_genre")),
    playlist_subgenre = any_of(c("subgenre", "playlist_subgenre")),
    soundcloud_link = any_of(c("soundcloud_link", "links"))
  ) %>%
  filter(!is.na(track), !is.na(artist), !is.na(popularity))
glimpse(SONGS)

```

Data ingest refers to the process of bringing raw data into an
analytical environment so that it can be examined, manipulated, and
prepared for statistical modeling. In this project, ingestion involved
importing a publicly accessible dataset from Hugging Face — a leading
repository for datasets commonly used in machine learning, artificial
intelligence, and data science. The dataset selected was appropriate
because it merged two distinct domains of listener behavior:
Spotify-style musical feature scoring and SoundCloud link-based
exposure, thus offering a hybrid perspective that reflects both
algorithmic recommendation and grassroots community engagement.

Within the R environment, the dataset was read using CSV import
functions and stored as a dataframe. Each row represented a unique track
streamed on SoundCloud, while columns captured metadata and analytical
properties of interest. Variables included track name, artist, album
title, external SoundCloud link, popularity score, playlist name,
playlist genre, playlist subgenre, and a set of Spotify-derived musical
features — such as danceability, tempo, energy, and valence. The dataset
originally consisted of 14,987 observations and 23 variables — a
sufficiently large sample to support meaningful statistical inference
and exploratory visualization.

The decision to use this dataset was also driven by its alignment with
the project's research goals. Because the study examines whether
playlist exposure influences streaming success, it is essential that the
dataset provides both playlist metadata and popularity metrics.
Additionally, the inclusion of musical features enables a secondary
layer of inquiry: whether popularity is driven by intrinsic qualities of
a song — such as mood or rhythm — or whether it is driven externally by
exposure mechanisms such as playlist placement.

Furthermore, the ingest stage required validating that the dataset could
be processed efficiently without extensive transformation. Basic checks
ensured that character fields were readable, numeric values were
correctly parsed, and variable types were appropriate for analyses such
as correlations, aggregations, and regression. With ingest complete, the
dataset was successfully loaded and prepared for subsequent cleaning and
restructuring steps.

# **Data Cleaning**

Now lets clean the data.

```{r}
PLAYLIST_TABLE <- SONGS %>%
  transmute(
    playlist_name   = playlist_name,
    artist_name     = artist,
    track_name      = track,
    album_name      = album,
    popularity      = popularity,
    playlist_genre  = playlist_genre,
    playlist_subgenre = playlist_subgenre,
    soundcloud_link = soundcloud_link
  )

glimpse(PLAYLIST_TABLE)





```

After the initial ingestion step, the dataset underwent a structured
data cleaning process designed to improve accuracy, enhance
interpretability, and prepare the data for meaningful exploration.
Because the goal of this project is to understand whether playlist
presence is associated with popularity, the cleaning stage prioritized
variables most directly related to playlist behavior, listener exposure,
and measurable popularity outcomes. The original dataset contained 23
raw variables — many of which were redundant, uninformative, or
irrelevant to the research question. These included ID hashes, file
directory information, metadata fields with excessive missingness, and
columns that offered no analytical value. Removing such noise was
necessary to create a more coherent, focused analytical structure.

The cleaned table ultimately retained **14,987 observations and 8 key
variables**, including playlist name, artist name, track name, album
name, popularity score, playlist genre, playlist subgenre, and the
corresponding SoundCloud link. These retained columns represent the
minimum required information needed to investigate playlist placement,
provide contextual metadata for songs, and link the data to an external
platform where tracks may be experienced. Selecting only the variables
that align directly with streaming dynamics ensures that subsequent
analyses remain interpretable and aligned to the central research
question.

Additionally, **tracks with missing popularity values were removed** to
ensure consistency across statistical outputs. Popularity is the
dependent variable in this study; therefore, missing values would
undermine the ability to compare songs fairly across playlists or within
clusters of musical attributes. Rather than attempting to impute missing
values — which would introduce artificial assumptions — the removal of
incomplete records ensures that results are derived entirely from real,
measurable data.

A key part of cleaning also included preparing the data for
playlist-level aggregation. Because the raw dataset recorded playlist
membership at a per-row level, individual tracks appeared multiple times
if they existed across several playlists. To evaluate playlist exposure
as a single measurable feature, a new aggregated track-level table was
created, counting the number of playlists each unique track–artist
combination appeared in. This restructuring makes playlist exposure
quantifiable and comparable across all songs.

This streamlined dataset — containing only the most analytically
relevant fields and complete popularity values — provides a balanced
combination of playlist context, track metadata, and platform linkage.
This clean foundation ensures that exploratory visualizations and
statistical models accurately reflect the true dynamics of how playlist
inclusion and musical characteristics relate to song popularity.

# **Data Exploration**

We define a "popular song" as one with a popularity that is greater than
or equal to 70.

```{r}
ppop_threshold <- 70
ppop_threshold

track_counts <- PLAYLIST_TABLE %>%
  distinct(playlist_name, track_name, artist_name, popularity) %>%
  count(track_name, artist_name, popularity, name = "playlist_appearances")

glimpse(track_counts)
```

To better understand how song popularity relates to playlist exposure, I
constructed a track-level summary table that aggregates playlist
information across the dataset. Each row in this table represents a
unique track and artist combination, along with the song’s popularity
score and the number of playlists in which it appears. Notably, most
tracks appear in only a single playlist regardless of their popularity
score, indicating that playlist inclusion in this dataset is relatively
sparse and not dominated by a small number of highly repeated songs.
This aggregation allows for direct comparison between popularity and
playlist appearances and serves as the foundation for subsequent visual
analyses examining whether more popular songs tend to receive greater
playlist exposure

## *Popularity vs Playlist Appearances*

```{r}

ggplot(track_counts, aes(x = popularity, y = playlist_appearances)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Popularity vs Playlist Appearances",
    x = "Track Popularity",
    y = "Number of Playlist Appearances"
  ) +
  theme_minimal(base_size = 13)



```

This scatter plot examines the relationship between track popularity and
the number of playlists in which a track appears. The majority of tracks
cluster at a single playlist appearance regardless of popularity score,
indicating that even highly popular songs are not necessarily included
in multiple playlists. While a small number of moderately to highly
popular tracks appear more frequently, the overall pattern shows no
strong upward trend. This suggests that playlist inclusion is not
strongly driven by popularity alone and may instead reflect playlist
curation strategies, genre specialization, or user preferences.

## ***Most danceable songs***

```{r}
SONGS %>%
  arrange(desc(danceability)) %>%
  select(track, artist, danceability, popularity, soundcloud_link) %>%
  slice_head(n = 5)


```

The table highlights the most danceable tracks in the dataset, with
danceability scores approaching the upper bound of the metric. Notably,
some of these tracks—such as *“ice ice baby”* and *“bad bad bad”*—also
exhibit high popularity, while others remain relatively less popular
despite their strong rhythmic characteristics. This reinforces the idea
that danceability contributes to popularity but does not guarantee
widespread success on its own

***Danceability vs Popularity***

```{r}
ggplot(SONGS, aes(x = danceability, y = popularity)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Danceability vs Popularity",
    x = "Danceability",
    y = "Popularity"
  ) +
  theme_minimal(base_size = 13)


```

This plot visualizes the relationship between danceability and
popularity across all tracks. A weak positive trend is visible,
indicating that tracks with higher danceability tend to be slightly more
popular on average. However, the substantial dispersion of points shows
that popularity varies widely at all danceability levels. This implies
that while danceability may enhance a track’s appeal, it is only one of
many factors influencing popularity.

## ***Energy Distribution of Popular Songs***

```{r}

SONGS %>%
  filter(popularity >= ppop_threshold) %>%
  ggplot(aes(x = energy)) +
  geom_histogram(bins = 30) +
  labs(
    title = paste0("Energy Distribution (Popularity ≥ ", ppop_threshold, ")"),
    x = "Energy",
    y = "Count"
  ) +
  theme_minimal(base_size = 13)

```

This histogram shows the distribution of energy levels among songs
classified as popular (popularity ≥ 70). Most popular tracks cluster in
the moderate-to-high energy range, particularly between 0.5 and 0.8,
indicating a preference for energetic but not extreme compositions. Very
low-energy tracks are relatively rare among popular songs, suggesting
that higher energy may be a common characteristic of broadly appealing
music.

***Valence (Mood) of Popular Songs***

```{r}
SONGS %>%
  filter(popularity >= ppop_threshold) %>%
  ggplot(aes(x = valence)) +
  geom_density(alpha = 0.5) +
  labs(
    title = paste0("Valence (Mood) Density (Popularity ≥ ", ppop_threshold, ")"),
    x = "Valence",
    y = "Density"
  ) +
  theme_minimal(base_size = 13)




```

This density plot illustrates the emotional tone of popular songs using
the valence measure. Popular tracks tend to have mid-range to moderately
high valence, indicating neutral to positive emotional moods. Extremely
low-valence (sad) songs appear less frequently among the most popular
tracks, suggesting that upbeat or emotionally balanced songs are more
common in popular playlists.

## ***Tempo vs Popularity***

```{r}

ggplot(SONGS, aes(x = tempo, y = popularity)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Tempo vs Popularity",
    x = "Tempo",
    y = "Popularity"
  ) +
  theme_minimal(base_size = 13)



```

The relationship between tempo and popularity appears weak and
non-linear, as shown by the relatively flat smoothed trend line.
Popularity remains fairly stable across a wide range of tempo values,
with no clear tempo range dominating popular songs. This indicates that
tempo alone does not play a major role in determining a track’s
popularity.

# **Statistical Analysis**

```{r}
# Correlation: Popularity vs Playlist Appearances
cor.test(track_counts$popularity, track_counts$playlist_appearances)

# Correlations with audio features
cor.test(SONGS$danceability, SONGS$popularity)
cor.test(SONGS$energy, SONGS$popularity)
cor.test(SONGS$tempo, SONGS$popularity)
cor.test(SONGS$valence, SONGS$popularity)

# Linear regression
model <- lm(popularity ~ danceability + energy + tempo + valence, data = SONGS)
summary(model)

```

Statistical analysis helps validate or refute patterns suggested by
exploratory visuals. In this project, multiple statistical techniques
were used to evaluate whether popularity is significantly associated
with playlist exposure and whether musical characteristics meaningfully
predict song success.

The first analysis performed was a **Pearson correlation test**, which
measured linear relationships between musical audio features and
popularity scores. Results showed that:

-   Danceability had a **weak but statistically significant** positive
    correlation (r ≈ 0.059), indicating that more danceable songs tend
    to be slightly more popular.

-   Energy demonstrated a **weak negative correlation** (r ≈ −0.091),
    suggesting overly intense songs may be less appealing to broad
    listener groups.

-   Tempo and valence showed **near-zero correlation**, suggesting no
    meaningful linear association.

However, correlation alone does not test combined predictive effects.
Therefore, a **multiple linear regression model** was applied, using
danceability, energy, tempo, and valence as predictor variables and
popularity as the outcome. While some coefficients were statistically
significant, the model’s **R² value was approximately 0.012**, meaning
that only **1.2% of variation in popularity** can be explained by
musical features alone.

This extremely low explanatory power means that — even when combined —
musical qualities account for **nearly none** of the factors that make
songs popular within this dataset. If intrinsic musical structure were
truly dominant, R² would be closer to 0.3 or above.

Next, playlist exposure was considered. Because most tracks appear in
only **one playlist**, playlist exposure lacks meaningful variance,
making correlation statistically undefined. In other words, playlist
count cannot explain popularity because playlist count itself rarely
exceeds one. Playlist presence within this dataset **does not appear to
materially influence** success.

Collectively, statistical tests strongly suggest that **popularity is
not determined by playlist count or musical composition**. Instead,
popularity is likely driven by external mechanisms such as social media
virality, artist influence, advertising campaigns, TikTok usage, and
algorithmic boost — none of which are represented in the provided
dataset.

# **Conclusion:**

This project set out to investigate whether playlist presence and
musical characteristics meaningfully influence song popularity across
music streaming platforms. Using a dataset containing nearly 15,000
SoundCloud-linked tracks enriched with Spotify-style audio features, the
analysis incorporated exploratory visualization and statistical testing
to evaluate how popularity is distributed and whether it aligns with
common assumptions about what makes a song successful.

Across the exploratory visualizations, the results consistently pointed
toward **very weak relationships** between musical characteristics and
popularity. For example, the **Danceability vs. Popularity scatterplot**
showed a cloud of data points with almost no visible trend, although a
slight positive slope on the fitted line suggests that more danceable
songs tend to have marginally higher popularity scores. However, the
density and heavy overlap of observations indicate that danceability
alone does not differentiate highly popular songs from those with very
low scores; popular songs span the entire danceability range.

The **Energy Distribution histogram**, filtered to include only tracks
with popularity ≥ 70, demonstrated that moderately energetic music
(values clustered between 0.5 and 0.8) appears most frequently among
popular tracks, suggesting energy may influence listener appeal. Yet
this distribution also revealed that extremely high-energy and
low-energy songs are both underrepresented among popular tracks —
meaning energy is associated with popularity, but only within a
constrained optimal middle range.

In contrast, the **Tempo vs. Popularity scatterplot** showed almost **no
relationship at all**, evidenced visually by a nearly flat smoothed
line. Popular songs can be slow, medium-tempo, or fast; this lack of
structure strongly suggests that tempo does not meaningfully shape
streaming success in this dataset.

Together, visual patterns align with statistical results showing that
musical features explain **only 1.2% of total popularity variance**,
confirming that intrinsic musical composition is not the dominant driver
of success. Additionally, playlist appearance data — not shown visually
due to sparse variation — revealed that most tracks appear in only one
playlist regardless of popularity, undermining the idea that playlist
exposure drives success through cumulative visibility.

Ultimately, this analysis concludes that **song popularity is only
weakly connected to measurable musical characteristics and not
meaningfully associated with playlist count** in this dataset. These
findings imply that digital music success is more likely driven by
**external factors** such as marketing, algorithmic placement, artist
fanbase size, repost-chain virality, and social trends on platforms like
TikTok — all of which fall outside the scope of this dataset and
represent a valuable direction for future research.
