---
title: "SoundCloud Analyis"
author: Hajara Muzammal
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    toc: true
    toc-depth: 3
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

# **Introduction:**

Music discovery increasingly occurs through playlists, where listeners
curate and share collections of songs across platforms such as
SoundCloud. Understanding which musical characteristics are associated
with popularity and playlist inclusion can provide insight into listener
preferences and emerging trends. This analysis focuses on tracks that
include SoundCloud links, allowing us to study how audio features,
sentiment, and popularity metrics relate to playlist usage.

Using a dataset of approximately 15,000 tracks with audio features,
playlist metadata, and SoundCloud URLs, this report explores the
relationship between popularity and musical characteristics such as
danceability, energy, tempo, and key. The goal is not to predict
popularity with a formal statistical model, but rather to visually and
descriptively identify patterns that distinguish popular songs from less
popular ones.

This work contributes to the broader question of how music spreads
across user-generated platforms and provides a data-driven perspective
on what makes songs more likely to appear in playlists.

Overarching Question:

**What factors influence the popularity of songs across major music
streaming platforms?**

My Question:

**Is track popularity associated with playlist inclusion?**

# **Data Ingest**

We use a publicly available dataset hosted on Hugging Face, which
contains playlist metadata, song characteristics, and direct SoundCloud
links.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

url <- "https://huggingface.co/datasets/Zuru7/Spotify_Songs_with_SoundCloud_links/resolve/main/song_df_normalised.csv"
SONGS_raw <- read_csv(url, show_col_types = FALSE)

# Standardize names (works even if you re-run the doc)
SONGS <- SONGS_raw %>%
  rename(
    track          = any_of(c("track", "track_name")),
    artist         = any_of(c("artist", "track_artist")),
    album          = any_of(c("album", "track_album_name")),
    popularity     = any_of(c("popularity", "track_popularity")),
    playlist_genre = any_of(c("genre", "playlist_genre")),
    playlist_subgenre = any_of(c("subgenre", "playlist_subgenre")),
    soundcloud_link = any_of(c("soundcloud_link", "links"))
  ) %>%
  filter(!is.na(track), !is.na(artist), !is.na(popularity))
glimpse(SONGS)

```

The dataset used in this analysis was obtained from Hugging Face and
contains Spotify song metadata linked to SoundCloud URLs. This dataset
is appropriate because it combines popularity metrics, playlist
information, and detailed audio features while also enabling direct
reference to SoundCloud content. The dataset consists of 14,987
observations and 23 variables, including track-level metadata, playlist
attributes, audio features, sentiment labels, and SoundCloud links,
making it well-suited for exploratory analysis of music popularity

# **Data Cleaning**

Now lets clean the data.

```{r}
PLAYLIST_TABLE <- SONGS %>%
  transmute(
    playlist_name   = playlist_name,
    artist_name     = artist,
    track_name      = track,
    album_name      = album,
    popularity      = popularity,
    playlist_genre  = playlist_genre,
    playlist_subgenre = playlist_subgenre,
    soundcloud_link = soundcloud_link
  )

glimpse(PLAYLIST_TABLE)





```

After initial ingestion, the dataset was cleaned to focus on variables
most relevant to playlist behavior and song popularity. The cleaned
table contains 14,987 observations and 8 key variables, including
playlist name, artist name, track name, album name, popularity score,
playlist genre and subgenre, and a corresponding SoundCloud link. Tracks
with missing popularity values were removed to ensure consistency across
analyses, allowing all visualizations and comparisons to rely on
complete and comparable observations. This streamlined structure
provides a balanced combination of playlist context, song metadata, and
platform linkage, making it well-suited for exploratory analysis of how
musical characteristics and playlist placement relate to popularity.

# **Data Exploration**

We define a "popular song" as one with a popularity that is greater than
or equal to 70.

```{r}
ppop_threshold <- 70
ppop_threshold

track_counts <- PLAYLIST_TABLE %>%
  distinct(playlist_name, track_name, artist_name, popularity) %>%
  count(track_name, artist_name, popularity, name = "playlist_appearances")

glimpse(track_counts)
```

To better understand how song popularity relates to playlist exposure, I
constructed a track-level summary table that aggregates playlist
information across the dataset. Each row in this table represents a
unique track and artist combination, along with the song’s popularity
score and the number of playlists in which it appears. Notably, most
tracks appear in only a single playlist regardless of their popularity
score, indicating that playlist inclusion in this dataset is relatively
sparse and not dominated by a small number of highly repeated songs.
This aggregation allows for direct comparison between popularity and
playlist appearances and serves as the foundation for subsequent visual
analyses examining whether more popular songs tend to receive greater
playlist exposure

## *Popularity vs Playlist Appearances*

```{r}

ggplot(track_counts, aes(x = popularity, y = playlist_appearances)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Popularity vs Playlist Appearances",
    x = "Track Popularity",
    y = "Number of Playlist Appearances"
  ) +
  theme_minimal(base_size = 13)



```

This plot shows the relationship between track popularity and the number
of playlist appearances in the dataset. All tracks appear exactly once
because the SoundCloud-linked dataset represents one playlist
association per track, meaning playlist frequency cannot vary; as a
result, no meaningful relationship between popularity and playlist
appearances can be inferred from this data, highlighting a structural
limitation of the dataset rather than a musical trend

## ***Most danceable songs***

```{r}
SONGS %>%
  arrange(desc(danceability)) %>%
  select(track, artist, danceability, popularity, soundcloud_link) %>%
  slice_head(n = 5)


```

This table highlights the five most danceable songs in the dataset,
ranked by Spotify’s danceability score. While these tracks all score
extremely high on danceability—indicating strong rhythm and suitability
for dancing—their popularity varies noticeably, ranging from moderately
popular to highly popular songs. This suggests that although
danceability contributes to a song’s appeal, it does not guarantee high
popularity on its own, reinforcing the idea that popularity is
influenced by multiple musical and contextual factors.

\## ***Danceability vs Popularity***

```{r}
ggplot(SONGS, aes(x = danceability, y = popularity)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Danceability vs Popularity",
    x = "Danceability",
    y = "Popularity"
  ) +
  theme_minimal(base_size = 13)


```

This scatter plot reveals a weak but positive relationship between
danceability and popularity. Although more danceable songs tend to have
slightly higher popularity on average, the wide dispersion of points
indicates that danceability is not a strong standalone predictor of a
song’s popularity.

## ***Energy Distribution of Popular Songs***

```{r}

SONGS %>%
  filter(popularity >= ppop_threshold) %>%
  ggplot(aes(x = energy)) +
  geom_histogram(bins = 30) +
  labs(
    title = paste0("Energy Distribution (Popularity ≥ ", ppop_threshold, ")"),
    x = "Energy",
    y = "Count"
  ) +
  theme_minimal(base_size = 13)

```

This histogram shows the distribution of energy levels among songs
classified as popular (popularity ≥ 70). Most popular tracks cluster in
the moderate-to-high energy range, particularly between 0.5 and 0.8,
indicating a preference for energetic but not extreme compositions. Very
low-energy tracks are relatively rare among popular songs, suggesting
that higher energy may be a common characteristic of broadly appealing
music.

\## ***Valence (Mood) of Popular Songs***

```{r}
SONGS %>%
  filter(popularity >= ppop_threshold) %>%
  ggplot(aes(x = valence)) +
  geom_density(alpha = 0.5) +
  labs(
    title = paste0("Valence (Mood) Density (Popularity ≥ ", ppop_threshold, ")"),
    x = "Valence",
    y = "Density"
  ) +
  theme_minimal(base_size = 13)




```

This density plot illustrates the emotional tone of popular songs using
the valence measure. Popular tracks tend to have mid-range to moderately
high valence, indicating neutral to positive emotional moods. Extremely
low-valence (sad) songs appear less frequently among the most popular
tracks, suggesting that upbeat or emotionally balanced songs are more
common in popular playlists.

## ***Tempo vs Popularity***

```{r}

ggplot(SONGS, aes(x = tempo, y = popularity)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Tempo vs Popularity",
    x = "Tempo",
    y = "Popularity"
  ) +
  theme_minimal(base_size = 13)



```

The relationship between tempo and popularity appears weak and
non-linear, as shown by the relatively flat smoothed trend line.
Popularity remains fairly stable across a wide range of tempo values,
with no clear tempo range dominating popular songs. This indicates that
tempo alone does not play a major role in determining a track’s
popularity.

# **Conclusion:**

This analysis examined whether track popularity is associated with
playlist inclusion in order to better understand broader factors that
influence song popularity across major music streaming platforms. Using
a SoundCloud-linked Spotify dataset, the results show that playlist
inclusion alone is not a strong driver of popularity: most tracks appear
in only one playlist regardless of their popularity score, indicating
that playlist presence does not scale proportionally with popularity.
While certain audio features such as danceability and energy exhibit
weak positive relationships with popularity—suggesting that rhythmically
engaging and moderately high-energy songs are more likely to achieve
higher popularity—these effects are modest and highly variable. Popular
songs also tend to cluster around mid-to-high valence levels, indicating
a preference for emotionally positive or balanced tracks rather than
extreme moods, while tempo shows little systematic relationship with
popularity. Taken together, these findings suggest that song popularity
is influenced by a combination of musical characteristics and platform
dynamics rather than simple playlist exposure alone, highlighting the
multifaceted nature of music consumption and discovery across streaming
platforms

# **Future Work**

There are several ways this project could be extended to more fully
address the factors that influence song popularity. First, incorporating
richer playlist data—such as playlist follower counts, track order
within playlists, and repeated appearances across multiple
playlists—would allow for a more nuanced analysis of how playlist
exposure impacts popularity. Second, integrating time-based information
(e.g., release year, time since release, or changes in popularity over
time) could help distinguish between long-term popularity and short-term
viral success. Additionally, expanding the dataset to include direct
SoundCloud engagement metrics, such as likes, reposts, or comments,
would strengthen cross-platform comparisons. Finally, future analyses
could move beyond exploratory visualization by incorporating predictive
models or resampling-based inference to quantify the relative importance
of different audio features and playlist dynamics, providing deeper
insight into the mechanisms that drive song popularity on streaming
platforms.
